CRUSE to Safe Cycling in Ireland

### 2023-10-03 paper.qmd
Add section on routing

The first version of the PCT was based on current and future potential uptake of cycling for single stage *travel to work* at desire line, zone, route, and route network levels [@lovelace2016].
It was launched in April 2017 as the government-endorsed tool for strategic cycle network planning, as part of the Cycling and Walking Investment Strategy [@cycling2017].
Extensions of the PCT approach have included estimation of benefits at the individual level [@woodcock2018], addition of travel to school network [@goodman2019], and improved modelling of impacts on health, environmental and distributional outcomes [@woodcock2021].
Initially developed just for England, the PCT was extended to cover all of Wales (for commuter data only) in 2018.

@@

- Domestic with cycle trips starting at home
- Domestic trips involving travel by other mode to the start of the cycle trip
- Cycle tourism, where start and end trips are different
- Sport cycling 
-->
```
## User interface {#sec-ui}

>>>>

The first version of the PCT was based on current and future potential uptake of cycling for single stage *travel to work* at desire line, zone, route, and route network levels [@lovelace2016].
It was launched in April 2017 as the government-endorsed tool for strategic cycle network planning and part of the Cycling and Walking Investment Strategy [@cycling2017].
Extensions of the PCT approach have included estimation of benefits at the individual level [@woodcock2018], addition of travel to school network [@goodman2019], and improved modelling of impacts on health, environmental and distributional outcomes [@woodcock2021].
Initially developed just for England, the PCT was extended to cover all of Wales (for commuter data only) in 2018.

@@

- Domestic with cycle trips starting at home
- Domestic trips involving travel by other mode to the start of the cycle trip
- Cycle tourism, where start and end trips are different
- Sport cycling 
-->
```

## Routing

Routes in CRUSE are generated by CycleStreets, a not-for profit transport consultancy and web development company who provide application programming interfaces (APIs) supplying a range of datasets for cycle planning and advocacy.
Although based in Cambridge, UK, CycleStreets provide routing services 'by cyclists, for cyclists' internationally, including in Ireland.
The CycleStreets routing engine is based on OpenStreetMap (OSM) data, which is continuously updated by a global community of volunteers.

CycleStreets' services can be used by anyone for free, but we commissioned a bespoke routing service for CRUSE for the following reasons:

- We needed to calculate hundreds-of-thousands of routes, which is against the terms of service of the free API.
- We needed to make changes to the routing profiles, including allowing routing on trunk roads, which are not permitted in the default routing profiles.
- We needed to be able to control the version of OSM data being used for the routing, enabling regular updates to the route networks as OSM data is updated.

## Scenarios of cycling potential {#sec-scenarios}

The datasets outlined above were used to generate estimates of cycling currently (the 'Baseline' scenario) and under four scenarios of cycling potential: Near Market, Climate Action Plan, Go Dutch, and Ebike.
Each of these scenarios is described below.
<!-- TODO: add table. -->

The Baseline scenario approximates current cycling levels.
As outlined

Near market
Climate Action Plan
Go Dutch
Ebike

## User interface {#sec-ui}
